---
title: "Primitive Pedestrian and Bike Crash Shiny"
output: flexdashboard::flex_dashboard
runtime: shiny
---

```{r}
library(flexdashboard)
library(shiny)
library(leaflet)
library(stringr)
library(tidyverse)
###############################
# Global Env.
######## Limitation; if you can get this to run on a SQL/Json query or using API, it would be best.
#setwd(dir = "C:/Users/b3nwi/OneDrive/Documents/data-exploration-ped-bike")
#myData <- read.csv(file = "Pedestrian_And_Bike_Crash_Data_-_Dataset.csv", header = TRUE)
myData <- read.csv(url("https://raw.githubusercontent.com/benjnguyen/pedestrian-bike-data/master/Pedestrian_And_Bike_Crash_Data_-_Dataset.csv"), header = TRUE)
# Some locations are empty in the data set
myData <- myData[-which(myData$Crash.Lat.Long.Location == ''), ]
dump = myData$Crash.Lat.Long.Location
myData_Locations <- as.vector(dump[which(dump != "")])
myData_Locations <- str_match_all(myData_Locations, "(?<=\\().+?(?=\\))")
myData_Locations <- as.numeric(unlist(str_extract_all(myData_Locations, "[\\.0-9e-]+")))
myData_LATITUDE <- c()
myData_LONGITUDE <- c()
for(i in 1:length(myData_Locations))
{
  if (i%%2==1)
  {
    myData_LATITUDE[i] <- myData_Locations[i]
  } else {
    myData_LONGITUDE[i] <- myData_Locations[i]
  }
}
myData$LATITUTDE <-  myData_LATITUDE[!is.na(myData_LATITUDE)]
myData$LONGITUDE <- myData_LONGITUDE[!is.na(myData_LONGITUDE)]
crashes <- c("Pedestrian", "Bike")
myData <- myData[which(myData$Crash.Type %in% crashes), ]
bikedata <- myData[which(myData$Crash.Type == 'Bike'), ]
peddata <- myData[which(myData$Crash.Type == 'Pedestrian'), ]
#tinyDataMatrix <- myData[sample(x = 1:nrow(myData), size = 100), ]

###############################
# Choices for crash-type
type <- c(
  "All Types" = "",
  "Pedestrian" = "PEDESTRIAN",
  "Bike" = "BICYCLE"
)
```


Column {.sidebar}
-----------------------------------------------------------------------
```{r}
################################
# UI
#ui <- fluidPage(
#  leafletOutput("map"),
#  p(),
#  radioButtons("vehicle", "Select crash type", choices = type, selected = "")
#)


fluidPage(
  helpText("Choose a crash type -- options are 'Bike' or 'Pedestrian'."),
  helpText("Choose the number of sites to appear on the map."),
  helpText("Toggle on clusters when number of sites is large to reduce clutter."),
  uiOutput("help1"),
  uiOutput("help2"),
  uiOutput("help3"),
  radioButtons("vehicle", "Select crash type", choices = type, selected = ""),
  numericInput("size", "Number of Sites", value = 100),
  checkboxInput("cluster", "Toggle Cluster", value = FALSE)
)
```



```{r}
output$help1<-renderUI({
  if(input$vehicle == ""){
     helpText("Maximum number of sites is", 
              nrow(myData), 
              "for the option 'All Types' \n")
  }
})

output$help2<-renderUI({
  if(input$vehicle == "PEDESTRIAN"){
    helpText("Maximum number of sites is", 
             nrow(myData[which(myData$Crash.Type == 'Pedestrian'), ]), 
             "for the option 'Pedestrian' \n")
  } 
})

output$help3<-renderUI({
  if(input$vehicle == "BICYCLE"){
     helpText("Maximum number of sites is", 
              nrow(myData[which(myData$Crash.Type == 'Bike'), ]), 
              "for the option 'Bike' \n")
  }
})

tinyDataMatrix <- reactive({
  if (input$vehicle == '') 
  {
    myData[sample(x = 1:nrow(myData), size = input$size), ]
  }
  else if (input$vehicle == 'BICYCLE') 
  {
    bikedata[sample(x = 1:nrow(bikedata), size = input$size), ]
  }
  else if (input$vehicle == 'PEDESTRIAN') 
  {
    peddata[sample(x = 1:nrow(peddata), size = input$size), ]
  }
})

#output$tinydf <- renderDataTable({
#  tinyDataMatrix() %>% 
#    select(Crash.Type, LONGITUDE, LATITUTDE) %>% 
#    head(5)
#})


# If cluster is true

output$map <- renderLeaflet({
  if (input$cluster) {
  leaflet() %>%
    addTiles() %>%
      addMarkers(lng = tinyDataMatrix()$LONGITUDE,
                 lat = tinyDataMatrix()$LATITUTDE,
                 popup = tinyDataMatrix()$Crash.Type,
                 clusterOptions = markerClusterOptions())
  } else {
  leaflet() %>%
    addTiles() %>%
      addMarkers(lng = tinyDataMatrix()$LONGITUDE,
                 lat = tinyDataMatrix()$LATITUTDE,
                 popup = tinyDataMatrix()$Crash.Type)
  }
})
  

```


Column
----------------------------------------
```{r}
dataTableOutput("tinydf")
leafletOutput("map")
```
  

### Background

This is a primitive version of an interactive dashboard that plots out crash site incidences of pedestrian and bike crashes. 

This data was provided by the St. Paul Police Department. The number of crash sites recorded at this time is 921; 613 of the crashes are of type 'Pedestrian', 300 are of type 'Bike', and the remaining are miscellaneously recorded values.

